---
title: "HW04 Part 1: Complete Chapter 3"
author: Kelsea Emmons
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Chapter 3

- Change "your name" in the YAML header above to your name.

### Section 3.1

Load the tidyverse packages. I've entered the first code chunk for you. **You must enter all subsequent code chunks and run the code.** 

```{r}
library("tidyverse")
```

If you get an error, then you did not install the `tidyverse` package like you were supposed to in the previous assignment. Go back to that assignment and complete that step. (This should not be the case, though.)

### Section 3.2: First steps

Enter your code chunks for Section 3.2 here.

This code pulls up the mpg data frame.
```{r}
ggplot2::mpg
```
This code tells me the dimensions of the table.
```{r}
dim(mpg)
```
This code pulls up the help page for 'mpg'.
```{r}
?mpg
```

This code creates a scatterplot of the class of the vehicle vs the drive train.
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = class, y = drv))
```

Get in the habit of entering a brief description of what each chunk of code does. Enter this above the chunk, like I did in Section 3.1.

#### Section 3.2 Questions

Answer the questions *completely.* Some answers will require you to write code in a code chunk and also to type text before or after the chunk. Other answers will require only code or only text. 

**1:** Run ggplot(data = mpg). What do you see?

When I ran this data chunk, it produced a table of cars and their respective specs.

**2:** How many rows are in `mpg`? How many columns? **Hint:** Use the `dim()` function. Type `dim()` in the console to learn what `dim()` does.

There are 234 rows and 11 columns in 'mpg'.

**3:** What does the `drv` variable describe? Read the help for `?mpg` to find out.

The 'drv' variable describes whether the vehicle is front-wheel drive, rear wheel drive, or four-wheel drive.

**4:** Make a scatterplot of `hwy` vs `cyl`.

This code chunk creates a scatterplot of the highway fuel efficiency vs the cars cylinder size.
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = hwy, y = cyl))
```


**5:** What happens if you make a scatterplot of `class` vs `drv`? Why is the plot not useful?

Whenever you plot 'class' vs 'drv', you end up with a plot that is pretty even across the board. Meaning, that the drive train within the vehicle is not necessarily dependent on the vehicle class, and therefore, rendering this scatterplot useless.

### Section 3.3: Aesthetic mappings


#### Section 3.3 questions

**1:** What’s gone wrong with this code? Why are the points not blue?

They needed to put the 'blue' level outside of the aes().

**2:** Which variables in `mpg` are categorical? Which variables are continuous? (*Hint:* type `?mpg` to read the documentation for the dataset). How can you see this information when you run mpg?

The variables 'model', 'cyl', 'year', 'trans', 'drv', 'fl', and 'class' are all categorical. The variables 'displ', 'cty', and 'hwy' are continuous variables.
When you run mpg you can see all the different categories.

**3:** Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables? **Note:** You need several code chunks here to answer this question.

Scatterplot of continuous variables with the color aesthetic.
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = cty, color = hwy))
```
Scatterplot of continuous variables with the size aesthetic.
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = cty, size = hwy))
```
Scatterplot of continuous variables with the shape aesthetic.
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = cty, shape = hwy))
```
The aesthetics only work with specific types of variables. For instance, when I tried to run the shape aesthetic with a continuous variable, it returned with an error. Shape aesthetic works best with categorical data and not continuous data because the amount of different shapes would be infinite.

**4:** What happens if you map the same variable to multiple aesthetics?

Scatterplot of the same variable with multiple aesthetics color and size.
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = cty, color = hwy, size = hwy))
```
If you map the same variable with multiple aesthetics, will get two legends for the same variable.

**5:** What does the stroke aesthetic do? What shapes does it work with? (*Hint:* use `?geom_point`)

The 'stroke' aesthetic modifys the width of the border of the shape. This works with shapes 0-14 and 21-24.

```{r}
?geom_point
```


**6:** What happens if you map an aesthetic to something other than a variable name, like `aes(colour = displ < 5)`?

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = class, y = drv, color = displ <5))
```

When you map an aesthetic to something other than a variable name, it creates a TRUE and FALSE legend.

### Section 3.5: Facets



#### Section 3.5 questions

**1:** What happens if you facet on a continuous variable? **Note:** Write a prediction, then make and run your code chunk.

I believe I will get an error because there could be and infinite number of facets.

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_wrap(~ displ, nrow = 2)
```



**2:** What do the empty cells in plot with `facet_grid(drv ~ cyl)` mean? How do they relate to this plot?

The empty plots mean that there are no data points that fit within those specific variables. In relation to the plot I just created, my plot has every grid filled, but the data is also continuous and not categorical.

**3:** What plots does the following code make? What does `.` do?

This code plots the single variable 'drv' in relation to the 'displ' variable. The `.` tells it to plot only in the rows dimension. (. in y-axis position)
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)
```

This code does the same as the last one except it tells it to plot in the columns dimension. (. in the x-axis position)
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)
```


**4:** Take the first faceted plot below. What are the advantages to using faceting instead of the colour aesthetic? What are the disadvantages? How might the balance change if you had a larger dataset?

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_wrap(~ class, nrow = 2)
```

With large groups of data like this, it is easier to read and decipher instead of different colors all acrossed a board. On the other hand, it is not as colorful and is kind of crowded. If you had an even larger dataset, it would become even more crowded and hard to differentiate.

**5:** Read `?facet_wrap`. What does `nrow` do? What does `ncol` do? What other options control the layout of the individual panels? Why doesn’t `facet_grid()` have nrow and ncol argument?

`nrow` tells the program how many rows to wrap the data into. `ncol` tells the program how many columns to wrap the data into. Other options to control the layout of the panels includes `strip.position`, `dir`, `drop`, `switch`, `as.table`, `labeller`, `shrink`, and `scales`. You wouldn't be able to use `nrow` and `ncol` with facet_grid() because there are more than one variables involved.


**6:** When using `facet_grid()` you should usually put the variable with more unique levels in the columns. Why?

You should put the variable with more unique levels or more "sporadic" data in the columns because it will be more spaced out over the grid instead of clustered.

### Section 3.6: Geometric objects



#### Section 3.6.1 Questions

**1:** What geom would you use to draw a line chart? A boxplot? A histogram? An area chart? **Note:** You will have to make some types of these charts in Part 2 of this assignment.

To draw a line chart, you would use the line geom. For boxplots, the boxplot geom. For histograms, use the histogram geom. To create an area chart, use the area geom.

**2:** Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions.

This code will create a scatterplot with `displ` and `hwy` as the x and y axis respectively, and have `drv` as different color points across the grid. There will also be a smooth line following the corrolation of the data.

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```


**3:** What does `show.legend = FALSE` do? What happens if you remove it?
**Note:** Skip the "Why do you think I used it earlier in the chapter?" question.

This hides the legend from view next to the grid. If you remove the `show.legend = FALSE` command, the legend will automatically be created and visible.

**4:** What does the `se` argument to `geom_smooth()` do?

The `se` argument asks the question if you would like to display confidence interval around smooth, which is set to TRUE by default.

**5:** Will these two graphs look different? Why/why not?

They will look the same. All that is different is that one code passed a set of mappings to ggplot(). Adding this infor to the original data and then adding the geoms just adds layers to what is already plotted. The other code will do the same thing, you would just have to type it and check it, twice.

**6:** Recreate the R code necessary to generate the following graphs. **Note:** Enter a separate code chunk for each graph, numbered 1 to 6, starting row-wise at upper left. 

1: upper left.

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, size = 3), show.legend = FALSE) +
  geom_smooth(mapping = aes(x = displ, y = hwy, size = 3), se = FALSE, show.legend = FALSE)

```


2: upper right.

```{r}
ggplot(data = mpg) + 
  geom_smooth(mapping = aes(x = displ, y = hwy, group = drv, size = 3), se = FALSE, show.legend = FALSE) +
  geom_point(mapping = aes(x = displ, y = hwy, size = 3), show.legend = FALSE)
```


3: middle left.

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv, size = 3)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
  

```


4: middle right.

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = drv, size = 3)) +
  geom_smooth(mapping = aes(x = displ, y = hwy, size = 3), se = FALSE, show.legend = FALSE)

```


5: lower left.

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = drv, size = 3)) +
  geom_smooth(mapping = aes(x = displ, y = hwy, group = drv, linetype = drv, size = 3), se = FALSE)

```


6: lower right. *Hint:* Read the help file for `geom_point`. Use `shape = 21`, `stroke = 2`, and `size = 3`.

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = drv, stroke = 2, size = 5), shape = 16, color = "white") +
  geom_point(mapping = aes(x = displ, y = hwy, color = drv, stroke = 2, size = 4), shape = 16)

```


### Section 3.7: Statistical transformations.

**Skip this section.** We may return to it later.


### Section 3.8 Position adjustments


#### Section 3.8 Questions

**1:** What is the problem with this plot? How could you improve it. *Hint:* Remember the size of the `mpg` data set from earlier in this exercise. Does the number of points seem correct? **Note:** Write the code to improve it. 

The problem with this plot is that some of the data points are overlapping each other. To fix it, you could use the `jitter` position.

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point(position = "jitter")
```


**2:** What parameters to `geom_jitter()` control the amount of jittering?

The `width` and `height` parameters control the amount of vertical and horizontal  jitter. 

**3:** Compare and contrast `geom_jitter()` with `geom_count()`.

The `geom_jitter` adds a small amount of random variation to the location of each point, while the `geom_count` counts the number of observations at each location, then maps the count to point area. Both are useful when you have discrete data and overplotting.

**4:** What is the default position adjustment for `geom_boxplot()`? Create a visualization of the `mpg` dataset that demonstrates it.

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_boxplot(aes(group = class))
```


### Section 3.9: Coordinate systems

**Note:** Read about and do only the code chunk associated with `coord_flip()`. `coord_flip()` is a great layer to use when you have categorical data with longish names as it makes the text easier to read.

```{r}
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + 
  geom_boxplot()
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + 
  geom_boxplot() +
  coord_flip()
```


### Section 3.10: The layered grammar of graphics

Just read this section as a contextual review of everything you just learned.


