---
title: "HW07 Part 1"
author: "Kelsea Emmons"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r}
library(tidyverse)
library(GGally)
library(patchwork)
```

###Part 1 - Recreate the Zuur Figures
####Import and Wrangle
```{r}
file_path <- "C:\\BI485\\emmons_kelsea\\hw07\\sparrows.txt"
sparrows_raw <- read_tsv("sparrows.txt", col_names = TRUE)
View(sparrows_raw)
sparrows <- sparrows_raw %>%
  mutate(box_group = "Sparrows",
         row_order = 1:nrow(.))
View(sparrows)
```

####Figure 2: Outliers
```{r}
plot_box <- sparrows %>% 
  ggplot(aes(x = box_group, 
             y = wingcrd)) +
  geom_boxplot(width = 0.25) +
  labs(x = NULL,
       y = "Wing length (mm)")
```


```{r}
plot_cleveland <- sparrows %>% 
  ggplot(aes(x = wingcrd,
             y = row_order)) +
  geom_point() +
  labs(x = "Wing length (mm)",
       y = "Order of the data")
```

```{r}
plot_box + plot_cleveland + coord_flip()
```

####Figure 3: Outliers
```{r}
p_culmen <- sparrows %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = culmen)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
p_nalospi <- sparrows %>%
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = nalospi)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
p_wt <- sparrows %>%
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = wt)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
p_wingcrd <- sparrows %>%
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = wingcrd)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
p_tarsus <- sparrows %>%
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = tarsus)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
p_head <- sparrows %>%
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = head)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
p_culmen + p_nalospi + p_wt + p_wingcrd + p_tarsus + p_head + plot_layout(ncol = 3)
```
Question: Which three variables appear to have outlier data?
  -The three variables that appear to have outlier date are nalospi, tarsus, and head variables.


####Figure 5: Distribution
```{r}
h1 <- sparrows %>%
  filter(Month %in% 6:8) %>% 
  ggplot() +
  geom_histogram(aes(x = wt),
                 binwidth = 0.5,
                 boundary = 0,
                 closed = "right",
                 color = "black") +
  scale_x_continuous(breaks = seq(14, 28, by = 2))
h1
```


```{r}
h2 <- sparrows %>%
  filter(Month %in% 6:8) %>% 
  mutate(Month = case_when(
    Month == 6 ~ "June",
    Month == 7 ~ "July",
    Month == 8 ~ "August")) %>% 
  ggplot() +
  geom_histogram(aes(x = wt),
                 binwidth = 0.5,
                 boundary = 0,
                 color = "black") +
  scale_x_continuous(breaks = seq(14, 28, by = 2)) +
  facet_wrap(~ Month, ncol = 1)
h2
```


```{r}
h1 + h2
```

####Figure 10: Relationships
```{r}
sparrows %>% ggscatmat(columns = 1:7)
```

####Figure 11: Interactions
```{r}
month_order = c("May", "June", "July", "August", "September")
sparrows %>% 
  filter(Month %in% 5:9,
         Sex != 0) %>% 
  mutate(Month = case_when(
    Month == 5 ~ "May",
    Month == 6 ~ "June",
    Month == 7 ~ "July",
    Month == 8 ~ "August",
    Month == 9 ~ "September"),
    Month = factor(Month, levels = month_order, ordered = TRUE),
    Sex = ifelse(Sex == 4, "Male", "Female")) %>% 
  ggplot(aes(x = wingcrd,
             y = wt)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  facet_grid(Sex ~ Month)
```

###Part 2: Habitat parameters for Aegla crabs
####Data Exploration: Crabs
```{r}
file_path <- "C:\\BI485\\emmons_kelsea\\hw07\\aegla_crabs.csv"
crabs <- read_csv("aegla_crabs.csv")
crabs <- subset(crabs, select = -c(Site,AT,WT,Shrimp,Aeglam,AeglaFe,AeglaFo))
crabs<- drop_na(crabs)
crabs <- crabs %>%
  mutate(row_order = 1:nrow(.))
View(crabs)
```

####Explore the Data
```{r}
p_width <- crabs %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = Width)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
p_Depth <- crabs %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = Depth)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
p_flow <- crabs %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = Flow)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
p_pH <- crabs %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = pH)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
p_TDS <- crabs %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = TDS)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
p_cond <- crabs %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = Cond)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
p_N <- crabs %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = N)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
p_Ni <- crabs %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = Ni)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
p_Nt <- crabs %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = Nt)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
p_phos <- crabs %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = Phos)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
p_Mg <- crabs %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = Mg)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
p_Ca <- crabs %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = Ca)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
p_Crab <- crabs %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = Crab)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

p_width + p_Depth + p_flow + p_pH + p_TDS + p_cond + p_N + p_Ni + p_Nt + p_phos + p_Mg + p_Ca + p_Crab + plot_layout(ncol = 6)
```
Depth, Flow, pH have outliers.
Depth because there is one value way up the chart for depth, well away from any other values.
Flow because the value is on the 0.0 line while no other values are.
pH because the value is lower than any other values and by itself.

```{r}
plot_flow <- crabs %>% 
  ggplot(aes(x = row_order, 
             y = Flow)) +
  geom_boxplot(width = 0.25) 
plot_flow
```

```{r}
plot_depth <- crabs %>% 
  ggplot(aes(x = row_order, 
             y = Depth)) +
  geom_boxplot(width = 0.25) 
plot_depth
```

```{r}
plot_pH <- crabs %>% 
  ggplot(aes(x = row_order, 
             y = pH)) +
  geom_boxplot(width = 0.25) 
plot_pH
```

I believe that Depth hold the actual true outlier. I believe this because after I plotted the variables I thought had outliers, you could actually see in the Flow and pH variables that there were multiple values outside of the norm, while in Depth, there was only one outlier.

####Cleveland Plots
```{r}
p_width <- crabs %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = Width)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
p_Depth <- crabs %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = Depth)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
p_flow <- crabs %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = Flow)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
p_pH <- crabs %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = pH)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
p_width + p_Depth + p_flow + p_pH + plot_layout(ncol = 2)
```

####Histograms
```{r}
h1TDS <- crabs %>%
  ggplot() +
  geom_histogram(aes(x = TDS),
                 binwidth = 1,
                 boundary = 0,
                 closed = "right",
                 color = "black") +
  scale_x_continuous(breaks = seq(0, 40, by = 5))
h1TDS
```

```{r}
h2Cond <- crabs %>%
  ggplot() +
  geom_histogram(aes(x = Cond),
                 binwidth = 2,
                 boundary = 0,
                 closed = "right",
                 color = "black") +
  scale_x_continuous(breaks = seq(2, 82, by = 4))
h2Cond
```

```{r}
h3Crab <- crabs %>%
  ggplot() +
  geom_histogram(aes(x = Crab),
                 binwidth = 1,
                 boundary = 0,
                 closed = "right",
                 color = "black") +
  scale_x_continuous(breaks = seq(0, 24, by = 2))
h3Crab
```

####Density Plots
```{r}
dp1TDS <- crabs %>% 
  ggplot() +
  geom_density(aes(x = TDS)) +
  scale_x_continuous(breaks = seq(0,40, by = 5))
dp1TDS
```

```{r}
dp2Cond <- crabs %>% 
  ggplot() +
  geom_density(aes(x = Cond)) +
  scale_x_continuous(breaks = seq(2,82, by = 4))
dp2Cond
```

```{r}
dp3Crab <- crabs %>% 
  ggplot() +
  geom_density(aes(x = Crab)) +
  scale_x_continuous(breaks = seq(0,24, by = 2))
dp3Crab
```
 
#### Patchwork
```{r}
histograms <- h1TDS + h2Cond + h3Crab + plot_layout(nrow = 3)
histograms 
```

```{r}
densityplots <- dp1TDS + dp2Cond + dp3Crab + plot_layout(nrow = 3)
densityplots
```
```{r}
histograms | densityplots 
```

####Explore Relationships Among All Variables
```{r}
crabs %>% ggpairs(columns = 1:14)
```

