---
title: "HW06 Part 2"
author: "Kelsea Emmons"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r}
library(tidyverse)
library(knitr)
library(smatr)
```

###Darters in Riffles

####Import Data
```{r}
file_path <- "C:\\BI485\\emmons_kelsea\\hw05\\darter.txt"
widths <- fwf_widths(c(3,11,1,1,1,3,4,4,4,3,3,2,2 ))
darter <- read_fwf("darter.txt", comment = "#", col_positions = widths)
col_names <- c("id","species","riffle","age","sex","length","depth","velocity","temperature","majsub","minsub","majtype","mintype")
colnames(darter) <- col_names
darter <- filter(darter, species == "spectabile" | species == "caeruleum" | species == "blennioides")
darter
```

####Wrangle Data
```{r}
darter <- mutate(darter,
       total_substrate = majsub + minsub)
```

####Change Riffle Values
```{r}
darter <- mutate(darter,
       riffle = ifelse(riffle == 1,
                           "Riffle 1",
                           "Riffle 2"))
```

Change Length Data from cm to mm.

```{r}
darter <- mutate(darter,
       length = length * 10)
```

Change the sex values of f and m to Female and Male.

```{r}
darter <- mutate(darter,
       sex = ifelse(sex == "f",
                           "Female",
                           "Male"))
```

####Summarize Data
```{r}
darter2 <- group_by(darter, species)
darter_summary <- darter2 %>%
  summarise(length = mean(length, na.rm = TRUE),
          depth = mean(depth, na.rm = TRUE),
          velocity = mean(velocity, na.rm = TRUE))
darter_summary
```

####Table of Summary Means
```{r}
darter_summary
```


####Plot 1
```{r}
ggplot(data = darter2, mapping = aes(x = species, y = length)) +
  stat_summary(
    fun.ymin = min,
    fun.ymax = max,
    fun.y = median) +
  geom_point(data = darter_summary, mapping = aes(x=species, y = length), shape = 15)  +
  coord_flip() +
  facet_wrap(~ riffle, nrow = 2) +
  xlab("Species") +
  ylab("Length (mm)")
  
```

####Plot 2
```{r}
ggplot(data = darter2, mapping = aes(x = species, y = length)) +
  stat_summary(
    fun.y = median, geom = "point", size = 2) +
  geom_point(data = darter2, mapping = aes(x=species, y = length), color = "gray50")  +
  stat_summary(
    fun.y = mean, geom = "point", size = 3, color = "maroon") +
  coord_flip() +
  facet_wrap(~ riffle, nrow = 2) +
  xlab("Species") +
  ylab("Length (mm)")
  
```





###Stress and Corticosterones in Birds
####Import Data
```{r}
"C:\\BI485\\emmons_kelsea\\hw06\\feathers.csv"
feathers <- read_csv("feathers.csv", locale = locale(decimal_mark = ","))
feathers
```
I believe that this data is tidy. Each variable has its own column, each observation has its own row and each value has its own cell.


```{r}
names(feathers) <- c("capture_date", "day", "site", "urban_score", "cortF", "cortB", "sex", "smi", "tarsus", "mass", "basal_cort", "stress_cort")

feathers
```

####Wrangle Data
```{r}
mean_tarsus <- mutate(feathers, mean_tarsus = mean(tarsus, na.rm = TRUE))
```
```{r}
mean_tarsus
```

```{r}
feathers <- mutate(feathers, log_mass = log(mass))

feathers <- mutate(feathers, log_tarsus = log(tarsus))

```

```{r}
major_axis <- sma(log_mass ~ log_tarsus, data = feathers)
summary(major_axis)
b <- coef(major_axis)[2]
```

```{r}
mean_tarsus <- mean(feathers$tarsus)
new_smi <- mutate(feathers,
                  new_smi = mass*mean_tarsus/tarsus^b)
new_smi
```

####Summary Data
```{r}
 feathers2 <- group_by(feathers, site)
basal_mean <- mean(feathers2$basal_cort, na.rm = TRUE)
stress_mean <- mean(feathers2$stress_cort, na.rm = TRUE)
basal_sd <- sd(feathers2$basal_cort, na.rm = TRUE)
stress_sd <- sd(feathers2$stress_cort, na.rm = TRUE)
basal_se <- mutate(feathers2,
                   basal_se = basal_sd/6.56)
stress_se <- mutate(feathers2,
                    stress_se = stress_sd/6.56)
           

```

####Graph Data
```{r}
ggplot(data = feathers) + 
  geom_boxplot(mapping = aes(x = site, y = new_smi))
```

```{r}
ggplot(data = feathers) +
  geom_point(mapping = aes(x = site, y = cortB)) +
  geom_errorbar(mapping = aes(x = site, ymin = basal_se - 1, ymax = basal_se +1))
```

```{r}
ggplot(data = feathers) +
  geom_point(mapping = aes(x = stress_mean, y = stress_se)) +
  geom_linerage(mapping = aes(x = stress_mean, ymin = stress_se -1, ymax = stress_se +1))
```

