---
title: "HW 06 Part 1: Complete the sections"
author: "Kelsea Emmons"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

- Change "your name" in the YAML header above to your name.

- As usual, enter the examples in code chunks and run them, unless told otherwise.

## Chapter 5: Data transformation

```{r}
library(tidyverse)
library(nycflights13)
```


Read [R4ds Chapter 5: Data transformation](https://r4ds.had.co.nz/transform.html), sections 5-7.

### Section 5.5: Add new variables with `mutate()`

Type and run all of the examples.

Pay careful attention to the useful creation functions in section 5.5.1. 

Do not answer any of the questions.

This code creates the columns gain and speed.
```{r}
flights_sml <- select(flights,
  year:day,
  ends_with("delay"),
  distance,
  air_time
  )
mutate(flights_sml,
  gain = dep_delay - arr_delay,
  speed = distance / air_time * 60
)
```

This code refers to the columns we just created.
```{r}
mutate(flights_sml,
       gain = dep_delay - arr_delay,
       hours = air_time/60,
       gain_per_hour = gain / hours
)
```

This code keeps only the new variables.
```{r}
transmute(flights,
          gain = dep_delay - arr_delay,
          hours = air_time / 60,
          gain_per_hour = gain / hours
)
```
This code breaks up the departure time into hour and minutes.

```{r}
transmute(flights,
          dep_time,
          hour = dep_time %/% 100,
          minute = dep_time %% 100
)
```

This code refers x to leading or lagging values.

```{r}
(x <- 1:10)
lag(x)
lead(x)
```

These codes run the cumulative sums and means of x.

```{r}
x
cumsum(x)
cummean(x)
```

These codes take y and rank them from min to max and vice versa.

```{r}
y <- c(1,2,2,NA, 3,4)
min_rank(y)
min_rank(desc(y))
```

Theses codes also order y in different ways.

```{r}
row_number(y)
dense_rank(y)
percent_rank(y)
cume_dist(y)
```

### Section 5.6: Grouped summaries with `summarise()`

Read and run all of the examples. Pay careful attention to the useful summary functions. Unless you really like baseball, skip the baseball part of 5.6.3.

This code collapses a data frame to a single row.
```{r}
summarise(flights, delay = mean(dep_delay, na.rm = TRUE))
```

This code groups the data frame by date.
```{r}
by_day <- group_by(flights, year, month, day)
summarise(by_day, delay = mean(dep_delay, na.rm = TRUE))
```

This code groups by destination and plots a grid of distance and delay.
```{r}
by_dest <- group_by(flights, dest)
delay <- summarise(by_dest,
                   count = n(),
                   dist = mean(distance, na.rm = TRUE),
                   delay = mean(arr_delay, na.rm = TRUE)
)
delay <- filter(delay, count > 20, dest != "HNL")
ggplot(data = delay, mapping = aes(x = dist, y = delay)) +
  geom_point(aes(size = count), alpha = 1/3) +
  geom_smooth(se = FALSE)
```

This code does the same thing as the one before but with the pipe.
```{r}
delays <- flights %>% 
  group_by(dest) %>% 
  summarise(
    count = n(),
    dist = mean(distance, na.rm = TRUE),
    delay = mean(arr_delay, na.rm = TRUE)
) %>% 
filter(count > 20, dest != "HNL")
```


This code groups the data set with missing values.
```{r}
flights %>% 
  group_by(year, month, day) %>% 
  summarise(mean = mean(dep_delay))
```


This code groups without the missing values.
```{r}
flights %>%
  group_by(year, month, day) %>% 
  summarise(mean = mean(dep_delay, na.rm = TRUE))
```

This code filters out the flights that are cancelled.
```{r}
not_cancelled <- flights %>% 
  filter(!is.na(dep_delay), !is.na(arr_delay))

not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(mean = mean(dep_delay))
```

This code takes the the not cancelled flights and groups them by tailnumber.
```{r}
delays <- not_cancelled %>% 
  group_by(tailnum) %>% 
  summarise(
    delay = mean(arr_delay)
  )

ggplot(data = delays, mapping = aes(x = delay)) +
  geom_freqpoly(binwidth = 10)
```
This code makes a scatterplot of the above data.

```{r}
delays <- not_cancelled %>% 
  group_by(tailnum) %>% 
  summarise(
    delay = mean(arr_delay, na.rm = TRUE),
    n = n()
  )

ggplot(data = delays, mapping = aes(x = n, y = delay)) +
  geom_point(alpha = 1/10)
```

This code filters out the groups with the smallest numbers of observations.
```{r}
delays %>% 
  filter(n > 25) %>% 
  ggplot(mapping = aes(x = n, y = delay)) +
  geom_point(alpha = 1/10)
```

This code shows the not cancelled flights and their average delay times.
```{r}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(
    avg_delay1 = mean(arr_delay),
    avg_delay2 = mean(arr_delay[arr_delay > 0])
  )
```

This code shows the standard deviations of distance.
```{r}
not_cancelled %>% 
  group_by(dest) %>% 
  summarise(distance_sd = sd(distance)) %>% 
  arrange(desc(distance_sd))
```

This code shows when the first and last flights leave each day.
```{r}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(
    first = min(dep_time),
    last = max(dep_time)
  )
```


This code shows the first and last departure time each day.
```{r}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(
    first_dep = first(dep_time),
    last_dep = last(dep_time)
  )
```

This code gives all variables with each obsevation in a separate row.

```{r}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  mutate(r = min_rank(desc(dep_time))) %>% 
  filter(r %in% range(r))
```

This code shows which destinations have the most carriers.

```{r}
not_cancelled %>% 
  group_by(dest) %>% 
  summarise(carriers = n_distinct(carrier)) %>% 
  arrange(desc(carriers))
```


This code shows just the counts of destinations.
```{r}
not_cancelled %>% 
  count(dest)
```

This code shows the counts of tailnumbers and the distance they flew.

```{r}
not_cancelled %>% 
  count(tailnum, wt = distance)
```

This code shows how many flights left before 5am. and if they were delayed.

```{r}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(n_early = sum(dep_time < 500))
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(hour_perc = mean(arr_delay > 60))
```


This code groups multiple variables per day.
```{r}
daily <- group_by(flights, year, month, day)
(per_day <- summarise(daily, flights = n()))
```
This does the same, but per month.
```{r}
(per_month <- summarise(per_day, flights = sum(flights)))
```

This does the smae, but per year.
```{r}
(per_year <- summarise(per_month, flights = sum(flights)))
```

This ungroups the data.
```{r}
daily %>% 
  ungroup() %>% 
  summarise(flights = n())

```

### Section 5.7: Grouped mutates (and filters)

Read and run all of the examples.
This code finds the worst members in each group.
```{r}
flights_sml %>% 
  group_by(year, month, day) %>% 
  filter(rank(desc(arr_delay)) < 10)
```

This code finds all groups bigger than a threshold.
```{r}
popular_dests <- flights %>% 
  group_by(dest) %>% 
  filter(n() > 365)
popular_dests
```

This code standardises to compute per group metrics.

```{r}
popular_dests %>% 
  filter(arr_delay > 0) %>% 
  mutate(prop_delay = arr_delay / sum(arr_delay)) %>% 
  select(year:day, dest, arr_delay, prop_delay)
```

